An Experimental Evaluation of Hybrid Querying on Vectors (EA\&B)

### 2025-07-31 Revision/response.tex

\noindent
\textbf{Response:} We believe this is an excellent suggestion. To provide a more comprehensive evaluation of the robustness of each algorithm, we introduce new experiments on two challenging multi-modal datasets.

First, we have incorporated a standard text-to-image dataset, which is widely adopted by existing multi-modal ANN algorithms. In this dataset, the base vectors are images while the query set consists of texts, creating a modality gap that demands stronger algorithm robustness to achieve good performance. Second, to further test performance on complex queries, we have constructed a more challenging dataset with a base set of 500,000 text vectors and 500,000 image vectors, and a query set containing 5,000 text and 5,000 image vectors. On this latter dataset, we perform both attribute filtering and range queries using combined text and image inputs to analyze the algorithm's robustness under more complex, mixed-modality conditions. The experimental results are shown in Figure~\ref{fig:attribute_multimodel} and Figure~\ref{fig:range_multimodel}.

As for complex query types that involve both attribute filtering and range queries simultaneously, to the best of our knowledge, existing algorithms—except for general-purpose systems such as Milvus and Faiss—do not support this functionality. For example, ACORN needs to build separate indexes for each scenario and cannot handle attribute filtering and range queries simultaneously. Therefore, we do not consider this type of complex query in our experiments but instead study the two query types separately. \textbf{(Please see Section 4.1.1, Section 4.2.3 and Secntion 4.3.3, highlighted in blue in the revised manuscript.)  
	%Please refer to the blue highlighted parts in  Section 4.1.1 paragraph 4 and  Section 4.2.4 and Section 4.3.4 under Multi-modal Queries.)
}

Section 4.1.1:\textcolor{blue}{
"Additionally, we use the Out-of-Distribution (OOD) dataset Text2Image and the large-scale dataset Deep100M to evaluate the generalization capability of the algorithms. The Text2Image dataset includes two versions. The original version contains 1 million image embeddings as the base dataset and uses 10 thousand text embeddings as the query set. The Text2Image-Mix version is composed of 0.5 million image embeddings and 0.5 million text embeddings as the base dataset, with a query set of 5 thousand image embeddings and 5 thousand text embeddings."
}

Section 4.2.3: \textcolor{blue}{
	"We conduct experiments on the Text2image-Mix dataset. This dataset contains image and text vectors and is a hybrid multi-modal dataset. As shown in Figure 9, UNG and NHQ consistently perform best, followed by CAPS. Although this dataset mixes two modalities, the vector space distributions of different modalities often differ significantly. In practice, this means that the experiments effectively treat the image and text subsets as separate datasets, since the ground truths for each query belong to the same modality. Therefore, the performance on this dataset is similar to the results shown in Figure 4. In contrast, the Text2image dataset shown in Figure 9 is different. The query set and the base dataset belong to two completely different modalities. Queries are far from the dataset points, and the ground truths are more dispersed. To find the ground truths, algorithms typically need to examine more points, which leads to degraded performance." 
}

Section 4.3.3:\textcolor{blue}{
		"Similar to attribute filtering, the performance of multi-modal range filtering queries shown in Figure 14 also aligns with the single-modality results in Figure 13. These algorithms on the Text2Image dataset exhibit a significant performance degradation. This is because these range query algorithms are almost entirely graph-based, and graph algorithms require more iterations to find true neighbors in OOD datasets."
}

### 2025-07-31 VLDB2026/main.tex

	\subsubsection{Attribute Filtering}
%	We summarize the applicability of different algorithms across various attribute filtering scenarios in Table~\ref{tab:algo_scenarios}. Puck and UNG excel on large-scale datasets (S1). ACORN-1 and UNG achieve fast index construction (S2). UNG performs well on different datasets (S3, S4). In addition, CAPS performs well on hard datasets with high LID (S3), while NHQ and StitchedVamana perform better on easy datasets with low LID (S4).
%	Faiss and CAPS can generate relatively compact indexes with low memory usage, which is suitable for resource-constrained situations (S6). Milvus and Faiss serve as general-purpose solutions (S5). UNG stands out for its higher recall, precision, and QPS, and performs well in the OOD scenario (S7, S8).
\textcolor{orange}{
Overall, UNG, NHQ, and FilteredDisk\\-ANN demonstrate strong performance. However, UNG suffers from performance degradation when the attribute count differs between indexing and querying. NHQ, due to its distance fusion design, also experiences performance drops when there is a significant mismatch in attribute numbers between indexing and querying, and its performance is sensitive to attribute distribution. }

@@

\textcolor{orange}{
CAPS, based on clustering, features low index space overhead. It offers flexible support for varying numbers of attributes and is optimized for low-selectivity attribute scenarios, but its overall performance is relatively moderate.
Puck is well-suited for large-scale datasets such as those with 100 million points. It offers fast index construction and low memory consumption in such scenarios. Moreover, it delivers high search performance in multi-threaded environments, making it an excellent choice for deployment on extremely large datasets.}

@@

	\textcolor{orange}{
	iRange has relatively low time and space costs for index construction; however, its search performance declines significantly with increasing query range. Despite offering threading parameters for indexing, it does not effectively utilize multi-core parallelism in practical scenarios.
	SeRF incurs low time and space overhead but achieves only moderate overall query performance, with notably weaker results in small-range query scenarios.
	DSG supports unordered dynamic insertions and range queries, but its index construction is limited to single-threaded execution, resulting in long build times and poor practical applicability.}

@@

\textcolor{orange}{
ACORN demonstrates average overall performance. Among its variants, ACORN-1 shows significant advantages in indexing speed on billion-scale datasets, but suffers from high memory consumption and delivers only moderate performance.}

\textcolor{orange}{
Faiss, Milvus, and VBASE handle complex scenarios with advanced filtering operations more effectively. While they offer strong flexibility, their overall search performance is relatively modest. Relatively speaking, PASE performs less well across various aspects.}

>>>>

	\subsubsection{Attribute Filtering}
%	We summarize the applicability of different algorithms across various attribute filtering scenarios in Table~\ref{tab:algo_scenarios}. Puck and UNG excel on large-scale datasets (S1). ACORN-1 and UNG achieve fast index construction (S2). UNG performs well on different datasets (S3, S4). In addition, CAPS performs well on hard datasets with high LID (S3), while NHQ and StitchedVamana perform better on easy datasets with low LID (S4).
%	Faiss and CAPS can generate relatively compact indexes with low memory usage, which is suitable for resource-constrained situations (S6). Milvus and Faiss serve as general-purpose solutions (S5). UNG stands out for its higher recall, precision, and QPS, and performs well in the OOD scenario (S7, S8).
\textcolor{orange}{
Overall, UNG, NHQ and FilteredDisk\\-ANN demonstrate strong performance. However, UNG suffers from performance degradation when the attribute count differs between indexing and querying. NHQ, due to its distance fusion design, also experiences performance drops when there is a significant mismatch in attribute numbers between indexing and querying, and its performance is sensitive to attribute distribution. }

@@

\textcolor{orange}{
CAPS, based on clustering, features low index space overhead. It offers flexible support for varying numbers of attributes and is optimized for low-selectivity attribute scenarios, but its overall performance is relatively moderate.
Puck is well-suited for large-scale datasets such as those with 100 million points. It offers fast index construction and low memory consumption in such scenarios. 
%Moreover, it delivers high search performance in multi-threaded environments, making it an excellent choice for deployment on extremely large datasets.
}

@@

	\textcolor{orange}{
	iRange has relatively low time and space costs for index construction. However, its search performance declines significantly with increasing query range. Despite offering threading parameters for indexing, it does not effectively utilize multi-core parallelism in practical scenarios.
	SeRF incurs low time and space overhead but achieves only moderate overall query performance, with notably weaker results in small-range query scenarios.
	DSG supports unordered dynamic insertions and range queries, but its index construction is limited to single-threaded execution, resulting in long build times and poor practical applicability.}

@@

\textcolor{orange}{
ACORN demonstrates average overall performance. Among its variants, ACORN-1 shows significant advantages in indexing speed on 100M-scale datasets, but suffers from high memory consumption and delivers only moderate performance.}

\textcolor{orange}{
Faiss, Milvus and VBASE effectively handle complex filtering by employing versatile "pre-filtering" or "post-filtering" strategies, granting them significant flexibility and broad applicability. This flexibility, however, is often counterbalanced by their relatively modest overall search performance. Relatively speaking, PASE performs less well across various aspects.}
